Chemistry of fossil Pinus pollen and the effect of acetolysis

### 2021-04-15 manuscript.Rmd
added Als comments; wrote pca bootstrap to test orientation and depths

The study of plant micro and macrofossil (spores, seed and pollen) is an integral part of Palaeoecology and Palaeobotany.
Pollen records are an important tool to reconstruct past floristic, vegetational and environmental changes [@magriNewScenarioQuaternary2006; @tzedakisVegetationChangeGlacialInterglacial1994; @huntleyMigrationSpeciesResponse1989; @noguesbravoCrackingCodeBiodiversity2018; @tinnerEcologyAbiesAlba2013].

@@

In recent years, the study of pollen grain chemistry has shown promising results for biological and palaeoecological research.
Fourier Transform Infrares (FTIR) spectroscopy has become a popular method for analysing pollen chemistry.
A number of FTIR studies have identified pollen chemical differences on genus level, which may extent to species level differences [@zimmermannChemicalCharacterizationIdentification2018; @jardineChemotaxonomyDomesticatedGrasses2019; @muthreichChemicalVariationsQuercus2020; @zimmermannInfraredSpectroscopyPollen2014; @depciuchIdentificationBirchPollen2018; @diehnDiscriminationGrassPollen2020]. 
This is a promising finding for the application of pollen identification with chemical methods in palaeoecological research, because it has the potential to increase taxomonic resolution @dellannaPollenDiscriminationClassification2009; @zimmermannAnalysisAllergenicPollen2016.
Further, chemical methods are suitable for automation and offer a more objective method than traditional morphological methods [ @mondolApplicationHighThroughputScreening2019; @zimmermannAnalysisAllergenicPollen2016; @guedesPollenRamanSpectra2014].
In addition to taxonomic differences in pollen chemistry, there is evidence of environmental signals that are recorded in the chemistry of pollen [@bagciogluMultiscaleVibrationalSpectroscopic2015; @kendelChemicalAnalysisPollen2020; @diehnCombiningChemicalInformation2020].
Several studies have shown links between solar and UV radiation and pollen chemistry as well as responses in pollen chemistry to temperature changes.

>>>>

The study of plant micro and macrofossil (spores, seed and pollen) is an integral part of palaeoecology and palaeobotany.
Pollen records are an important tool to reconstruct past floristic, vegetational and environmental changes [@magriNewScenarioQuaternary2006; @tzedakisVegetationChangeGlacialInterglacial1994; @huntleyMigrationSpeciesResponse1989; @noguesbravoCrackingCodeBiodiversity2018; @tinnerEcologyAbiesAlba2013] and are important tools to address ecological processes with timescales that operate on centennial/decadal timescales.

@@

In recent years, the study of pollen-grain chemistry has shown promising results for biological and palaeoecological research.
Fourier Transform Infrared (FTIR) spectroscopy has become a popular method  for characterising the chemical variation within pollen and spores.
A number of FTIR studies have identified pollen chemical differences on genus level, which may extent to species level differences [@zimmermannChemicalCharacterizationIdentification2018; @jardineChemotaxonomyDomesticatedGrasses2019; @muthreichChemicalVariationsQuercus2020; @zimmermannInfraredSpectroscopyPollen2014; @depciuchIdentificationBirchPollen2018; @diehnDiscriminationGrassPollen2020]. 
This is a promising finding for the application of pollen identification with chemical methods in palaeoecological research, because it has the potential to increase taxomonic resolution @dellannaPollenDiscriminationClassification2009; @zimmermannAnalysisAllergenicPollen2016.
Further, chemical methods are suitable for automation and offer a more objective method than traditional morphological methods [ @mondolApplicationHighThroughputScreening2019; @zimmermannAnalysisAllergenicPollen2016; @guedesPollenRamanSpectra2014].
In addition to taxonomic differences in pollen chemistry, there is evidence of environmental signals that are recorded in the chemistry of pollen [@bagciogluMultiscaleVibrationalSpectroscopic2015; @kendelChemicalAnalysisPollen2020; @diehnCombiningChemicalInformation2020].
Several studies have shown links between solar and UV radiation and pollen chemistry as well as responses in pollen chemistry to temperature changes.
